A fully labelled proof system for intuitionistic modal logics

### 2020-09-01 FullyLabelled/JLC/system.tex
sec. 3-5

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
%\sonia{This is an interesting alternative view but for me the starting point really is the Maffezioli, Naibo and Negri paper. Maybe rephrase slightly to acknowledge their work too.}
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{str:2017maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this naive translation would also yield the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A \IMP B} \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, \accs xz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, \accs xz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{str:2017maehara}, these are the two rules that force
single-conclusion. 
%This is the reason why in our system we need the
%$\le$ in the relational atoms; we need to go to the future of the
%current world.  
%
In our system, this phenomenon is replaced by a re-positioning of the considered formulas to a fresh label.
%
In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that make use of the
pre-order relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.
%
These can be obtained by applying the axioms-as-rules methodology as in~\cite{maffezioli:etal:synthese13}.

@@

>>>>

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
%\sonia{This is an interesting alternative view but for me the starting point really is the Maffezioli, Naibo and Negri paper. Maybe rephrase slightly to acknowledge their work too.}
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{str:2017maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this naive translation would also yield the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A \IMP B} \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, \accs xz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, \accs xz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{str:2017maehara}, these are the two rules that force
single-conclusion. 
%This is the reason why in our system we need the
%$\le$ in the relational atoms; we need to go to the future of the
%current world.  
%
In our system, this phenomenon is replaced by a re-positioning of the considered formulas to a fresh label.
%
In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that make use of the
pre-order relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.\footnote{These are the same as for {\bf G3I} as presented in~\cite{DN12}.\todo{Dyckhoff,Negri:Proof Analysis in Intermediate Logics}}

%
%%These can be obtained by applying the axioms-as-rules methodology as in~\cite{maffezioli:etal:synthese13}.

@@

\begin{remark}
  As mentioned above, the monotonicity rules $\llabrn{mon}$ and $\rlabrn{mon}$ are a form of contraction. So, it is not a surprise that the contraction rules
  \begin{equation*}
    \vlinf{\llabrn{cont}}{}{\B, \Left, \labels{x}{A} \SEQ \Right}{\B, \Left, \labels{x}{A}, \labels{x}{A} \SEQ \Right}
    \hskip8em
    \vlinf{\rlabrn{cont}}{}{\B, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{A}}
  \end{equation*}
  are admissible in our system, as they are derivable as follows:
  \begin{equation*}
    \vlderivation{
      \vlin{\rn{refl}}{}
	   {\B, \Left, \labels{x}{A} \SEQ \Right}
	   {\vlin{\llabrn{mon}}{}
	     {\B, \futs{x}{x}, \Left, \labels{x}{A} \SEQ \Right}
	     {\vlhy{\B, \futs{x}{x}, \Left, \labels{x}{A}, \labels{x}{A} \SEQ \Right}}}
    }
    \hskip6em
      \vlderivation{
      \vlin{\rn{refl}}{}
	   {\B, \Left \SEQ \Right, \labels{x}{A}}
	   {\vlin{\rlabrn{mon}}{}
	     {\B, \futs{x}{x}, \Left \SEQ \Right, \labels{x}{A}}
	     {\vlhy{\B, \futs{x}{x}, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{A}}}}
    }

  \end{equation*}

\review{the authors of the paper under review neglect an important caveat in the design of rules of contraction-free labelled sequent calculi: to guarantee that contraction be admissible one needs to ensure that a certain closure condition be satisfied, see:

\emph{• Negri, S. Proof analysis in modal logic. Journal of Philosophical Logic, 34, 507–544, 2005.}}\lutz{I think the remark above takes care of that.}

%\marianela{Should we mention Negri's paper here? }

%\marianela{In the same way as we did with monotonicity, trying to avoid the problems that contraction has in a cut elimination proof, it is preferable to have a system in which this rule is admissible.
%	}

\review{Also the proofs of the structural properties closely follow the methodology of the paper by Maffezioli, Naibo \& Negri. All this has to be said expressis verbis, and appropriate citations have to be added.}
%\marianela{we are not saying that in the next paragraph?}
\lutz{I just checkt again, and I refuse to do that. I see two ways to proceed: Either we just ignore the comment, or we write in the answer to the reviewer that the whole procedure is standard (and we can even cite in the paper the Troelstra-Schwichtenberg Textbook). We do cite that paper somewher, but they should not get more credit than they deserve.}
